---
output:
  html_document: default
  pdf_document: default
---

# Overview/Executive Summary


This project has as scope to create a movie recommendation system using the MovieLens dataset. There are two versions of this movielens: one included in the dslabs package is just a small subset of a much larger dataset and the second it is the dataset with millions of ratings, called Movielens 10M. This project it is going to work with the last one, this add the difficulty of more data that required more time y computing resources but it gives the opportunity to work with more realistic big data.

This prediction systems will be validated using part of this dataset, part that is not used in the training.

Recommendation systems are referred to as systems and tools that provide suggestions for the items the user uses.
Recommendation systems are more complicated machine learning challenges because each outcome has a different set of predictors. For example, different users rate a different number of movies and rate the movies in different ways.
The inputs to these systems are item’s data, user profile and most importantly, the behavior of each user's access to the items, which is referred to as Usage. System outputs is a collection of products and items that the user will mostly like or buy.



...You can find the entire latest MovieLens dataset here. You will be creating your own recommendation system using all the tools we have shown you throughout the courses in this series. We will use the 10M version of the MovieLens dataset to make the computation a little easier. ...



The MovieLens dataset is a table witch each row it is rating to a movie provided by one user and has the following structure:

```{r initial1, include = FALSE}
library(dslabs)
library(tidyverse)
library(rafalib)
data("movielens")

```


```{r initial2, echo=TRUE}
names(movielens)
```

Looking some rows (for space it is shown in two parts)
```{r initial3, echo=FALSE}
movielens_head <- head(movielens)
movielens_head %>% select(movieId,title, year)
movielens_head %>% select(genres,userId, rating, timestamp)
```


## Evaluation
The rating is an numeric running from 0.5 througth 5 by 0.5 step. This menas 10 possible options to the user who make the rating.
For evaluate the algorithm we are going to use the Root Mean Square Error (RMSE)[https://en.wikipedia.org/wiki/Root-mean-square_deviation]. This take avantage of the numeric characteristic of rating and in R can be represented as:

```{r initial4, echo=TRUE}
RMSE <- function(true_ratings, predicted_ratings){
  sqrt(mean((true_ratings - predicted_ratings)^2))
}
```

## Approach
The develop of this predicting rating system for moveis can be explained in the follows steps:

1.- Start creating the model and explained why this approach was taken. The main, but not the only, approach is in base of Data Science: Machine Learning from	Harvard University (course HarvardX+PH125.8x). The objective of the final model it is to get and get  a RSME less than xxxx in the validation set.

The algorithm is developed using the training (edx) set. For a final test of the final algorithm, predict movie ratings in the validation set (the final hold-out test set) as if they were unknown. In others words, the validation data (the final hold-out test set) should NOT be used for training, developing, or selecting your algorithm and it should ONLY be used for evaluating the RMSE of your final algorithm

2.- Present the algorithm, train with the train data, adjust the parameter and concluded with a tuned machine

3.- Continue with the final validation with this trained algorithm 

            that describes the dataset and summarizes the goal of the project and key steps that were performed
            
