---
output:
  html_document: default
  pdf_document: default
---

# Overview/Executive Summary

                  In this section we describe the dataset and summarizes the goal of the project and key steps that were performed
      
This project is part of the Capstone of the Professional Certificate Program of Data Science and it has as scope to create a movie recommendation system using the MovieLens dataset. There are several versions of this movielens, for example the one included in the dslabs package is just a small subset of a much larger dataset and the one we are going to use with a dataset with 10 millions of ratings, called Movielens 10M. This big dataset add the difficulty of more data require more time y computing resources, that force to discard some high consume resource algorithm because it can not ran in R with this volume but it gives the opportunity to work with more realistic big data.

This prediction systems will be validated using part of this dataset, part that is not used in the training.

Recommendation systems are referred to as systems and tools that provide suggestions for the items, in this case a movie, the user uses. Recommendation systems are more complicated machine learning challenges because each outcome has a different set of predictors. For example, different users rate a different number of movies and rate the movies in different ways. The inputs to these systems are item’s data, user profile and most importantly, the behavior of each user's access to the items, which in our case it is the rating. System outputs is a collection of products and items that the user will mostly like.





The MovieLens dataset is a table witch each row it is rating to a movie provided by one user and has the following structure:

```{r initial, include = FALSE}
##########################################################
# Create edx set, validation set (final hold-out test set)
##########################################################

# Note: this process could take a couple of minutes

if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(caret)
library(data.table)

# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")

# if using R 3.6 or earlier:
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                            title = as.character(title),
                                            genres = as.character(genres))
# if using R 4.0 or later:
# movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(movieId),
#                                            title = as.character(title),
#                                            genres = as.character(genres))


movielens <- left_join(ratings, movies, by = "movieId")

# Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding") # if using R 3.5 or earlier, use `set.seed(1)`
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% 
      semi_join(edx, by = "movieId") %>%
      semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

rm(dl, ratings, movies, test_index, temp, movielens, removed)
```


```{r initial2, echo=TRUE}
names(edx)
```

Looking some rows (for space it is shown in two parts)
```{r initial3, echo=FALSE}
head(edx)

```


## Evaluation
The rating is an numeric running from 0.5 througth 5 by 0.5 step. This means 10 possible options to the user who make the rating.
For evaluate the algorithm we are going to use the Root Mean Square Error (RMSE) [https://en.wikipedia.org/wiki/Root-mean-square_deviation]. This take advantage of the numeric characteristic of rating and in R can be represented as:

```{r initial4, echo=TRUE}
RMSE <- function(true_ratings, predicted_ratings){
  sqrt(mean((true_ratings - predicted_ratings)^2))
}
```

## Approach
The develop of this predicting rating system for moveis can be explained in the follows steps:

1.- Start creating the model and explained why this approach was taken. The main, but not the only, approach is in base of Data Science: Machine Learning from	Harvard University (course HarvardX+PH125.8x). The objective of the final model it is to get and get a *RMSE < 0.86490* in the validation set.

The algorithm is developed using the training (edx) set. For a final test of the final algorithm, predict movie ratings in the validation set (the final hold-out test set) as if they were unknown. In others words, the validation data (the final hold-out test set) should NOT be used for training, developing, or selecting your algorithm and it should ONLY be used for evaluating the RMSE of your final algorithm

2.- Present the algorithm, train with the train data, adjust the parameter and concluded with a tuned machine

3.- Continue with the final validation with this trained algorithm 

            that describes the dataset and summarizes the goal of the project and key steps that were performed
            
\newpage